Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-19 Manuscript/manuscript.tex

\subsection{Salt concentrations and binding affinity}
Experimental works providing the lipid head group order
parameters report the salt concentrations in two different ways.
For DPPC, salt concentrations before solvating
the lipids are used~\cite{akutsu81}. Instead, for POPC,
the salt concentration of supernatant after solvation is
measured using atomic absorption~\cite{altenbach84}.
In this work, we will use the latter definition. For this we
will measure the salt concentration in the water bulk region,
\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
Note that in our previous study, Ref.~\citenum{catte16},
we calculated the concentration as in the DPPC experiments.
Despite the two concentrations may differ significantly for CaCl$_2$ systems
these differences do not affect qualitatively the conclusions neither in this
nor in our previous work~\cite{catte16}.

To quantify the ion binding affinity to a membrane, we
calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
This quantity does not depend on the definition of the interfacial region
and the position of the Gibbs dividing plane. Therefore, we consider our whole simulation box as
an interface flanked by a water phase and an ideal phase with the characteristics of to
the hydrophobic region in the middle of the membrane where no ions or waters are present.
This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where $n_w$ and $n_i$ are the total number of waters and ions in the system;
$C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
and $A_b$ is the size of the box in the membrane plane. Therefore, this quantity
does not depend on any arbitrary definition of a bound ion to a membrane and
therefore allow us to accurately compare between different models. Note that
because bilayers have an interface in both leaflefts, the total area of the
interface is twice the area of the membrane, $2A_b$.

@@

\vspace{0.5cm}
\chemfig{
\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  (-[:0,3.5,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) 
} .
\vspace{0.5cm}

AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35 or 50 surfactants per leaflet.
The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
First 20~ns were omitted from the analysis.

@@

The binding affinity of aqueous cations in lipid 
bilayers can be measured by using the headgroup 
order parameters, because they decrease monotonously  
with the bound positive charge~\cite{seelig87,catte16}. 
The headgroup order parameter responses 
to aqueous NaCl and CaCl$_2$ concentrations 
are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}
from experiments (DPPC \cite{akutsu81} 
and POPC \cite{altenbach84}) and different simulation models for POPC.

>>>>

\subsection{Salt concentrations and binding affinity}
Experimental works providing the lipid head group order
parameters report the salt concentrations in two different ways.
For DPPC, salt concentrations before solvating
the lipids are used~\cite{akutsu81}. Instead, for POPC,
the salt concentration of supernatant after solvation is
measured using atomic absorption~\cite{altenbach84}.
In this work, we will use the latter definition. For this we
will measure the salt concentration in the water bulk region,
\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
Note that in the previous study, Ref.~\citenum{catte16},
the concentration was calculated as in the DPPC experiments.
Despite of the measurable differences between two concentrations for CaCl$_2$ systems,
the qualitatively conclusions are not affected neither in this
nor in our previous work~\cite{catte16}.

To quantify the ion binding affinity to a membrane, we
calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
This quantity can be used to compare binding affinity between different models,
because it does not depend on the definition of the interfacial region
or the position of the Gibbs dividing plane between two bulk regions.
Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
as a bulk region, which is separated by a monolayer interface from 
a another region with bulk concentration of ions in water.
Since the boundary of the latter region is set to the edge of simulation box,
all molecules in simulation are within the defined interfacial region. 
%our whole simulation box as
%an interface flanked by a water phase and an ideal phase with the characteristics of to
%the hydrophobic region in the middle of the membrane where no ions or waters are present.
This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where $n_w$ and $n_i$ are the total number of waters and ions in the system;
$C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
and $A_b$ is the size of the box in the membrane plane.
The total area of the interface is twice the area of the membrane, $2A_b$,
because bilayers have an interface in both leaflefts.

@@

\vspace{0.5cm}
\chemfig{
(-[:0,1.0,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  
} .
\vspace{0.5cm}

AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35 or 50 surfactants per leaflet. Parameters for Bromium ions were
not available in the standard Gromacs files for Amber force field, thus the
parameters for Chloride ions were used instead.
The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
First 20~ns were omitted from the analysis.

@@

The binding affinity of aqueous cations in lipid 
bilayers can be measured by using the headgroup 
order parameters, because they decrease proportionally  
to the bound positive charge~\cite{seelig87,catte16}. 
The headgroup order parameter responses 
to aqueous NaCl and CaCl$_2$ concentrations 
are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}
from experiments (DPPC \cite{akutsu81} 
and POPC \cite{altenbach84}) and different simulation models for POPC.
